---
title: "Project 3"
author: "Rachel Fellman & Sabrina Dahl"
params: 
  Ed: 1
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(caret)
```

# Introduction

# Data

We will start by reading in the data with the `read_csv` function with a relative path. 

```{r, message=FALSE}
diabetes <- read_csv("diabetes_binary_health_indicators_BRFSS2015.csv")
```

Next we will combine the 1 and 2 groups in the `Education` variable into a new group. We will do this using the `mutate` and `ifelse` functions. The new value for the 1 and 2 groups will just be 1, so `Education` can equal 1, 3, 4, 5, or 6.

```{r}
diabetes.1 <- diabetes %>% 
  mutate(Education = ifelse(Education < 3, 1, Education))
```

# Summarizations

We will start by doing some exploratory data analysis on the full data set. 

Since we will be creating models for different education levels, we will split up the data into smaller data set for each of the 5 education levels to do EDA. We will do this using our parameters
```{r}
#create subsetted data set for each level of education using the params.
diabetes.ed <- diabetes.1 %>% 
  filter(Education == params$Ed)
```

# Modeling

Before doing any modeling we will split the data into a training and test set using the `createDataPartition` function from the `caret` package. The test/train data will be split with a 70/30 ratio.

```{r}
#make things reproducible with set.seed (I have no idea what number we are supposed to put here because I don't really know what this function does)
set.seed(90)
#create index
train.index <- createDataPartition(diabetes.ed$Diabetes_binary, p = .7, list = FALSE)
#create train and test sets
diabetes.train <- diabetes.ed[train.index, ]
diabetes.test <- diabetes.ed[-train.index, ]
```


# Model Selection

