---
title: "Project 3"
author: "Rachel Fellman & Sabrina Dahl"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(caret)
```
test test does this commit to my branch only 
# Introduction

# Data

We will start by reading in the data with the `read_csv` function with a relative path. 

```{r, message=FALSE}
diabetes <- read_csv("diabetes_binary_health_indicators_BRFSS2015.csv")
```

Next we will combine the 1 and 2 groups in the `Education` variable into a new group. We will do this using the `mutate` and `ifelse` functions. The new value for the 1 and 2 groups will just be 1, so `Education` can equal 1, 3, 4, 5, or 6.

```{r}
diabetes.1 <- diabetes %>% 
  mutate(Education = ifelse(Education < 3, 1, Education))
```

# Summarizations

We will start by doing some exploratory data analysis on the full data set. 

# Modeling

Before doing any modeling we will split the data into a training and test set using the `createDataPartition` function from the `caret` package. The test/train data will be split with a 70/30 ratio.

```{r}
#make things reproducible with set.seed (I have no idea what number we are supposed to put here because I don't really know what this function does)
set.seed(90)
#create index
train.index <- createDataPartition(diabetes.1$Diabetes_binary, p = .7, list = FALSE)
#create train and test sets
diabetes.train <- diabetes.1[train.index, ]
diabetes.test <- diabetes.1[-train.index, ]
```


# Model Selection

