---
title: "ST558 Project 3"
author: "Rachel Fellman & Sabrina Dahl"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(purrr)
```


# Description
This repository is where Sabrina Dahl and Rachel Fellman will be working together on project 3 for ST 558. We will analyze data using different models and create 5 output files that will be linked below.

# R Packages
`tidyverse`
`caret`
`rmarkdown`
'pwalk`

# Code to Render Files

To render the files we first have to create a data frame that has the file names to output to and a list with each name for using in the `render` function. 
```{r, eval =FALSE}
#get education levels
ed.levels <- unique(diabetes.1$Education)
#create filenames
output_file <- paste0(ed.levels, ".md")
#create a list which just the education level parameter
params = lapply(ed.level, FUN = function(x){list(ed = x)})
#put into data frame
reports <- tibble(output_file, params)
```

Next, using the `pwalk` function, we will knit the documents.
```{r, eval = FALSE}
pwalk(reports, render, input = "project3.rmd")
```



# Links to files
